---
title: "Statistical Inference Assignment - Part 1"
author: "Tomás A. Maccor"
date: "23/3/2020"
output: pdf_document
---

```{r setup, include=FALSE, echo = FALSE}
library(ggplot2)

```

## Synopsis

In this assigment we are going to verify if the postulates of the Central Limit Theorem (CLT) stand for an exponential distribution density. That is, as we obtain a higher **N** of random samples from an an Exponention distribution, the distribution of the mean of each of these N random sampless will approximate a Normal distribution (regardless of the fact that the samples are obtained from an exponential distribution density). 

We will also try to prove another of the CLT postulates: that the average of the sample means will be the population mean. In other words, if we add up the means from all of the samples & calculate the average ----> that average will be the actual population mean. The same can be applied for finding the standard deviation of your population (distribution, in this case)

## Basic Facts

### EXPONENTIAL distribution parameters
Mean =  1/\( \lambda \)    
\( \lambda \) (lambda) = rate parameter for the distribution
St. dev. = 1/\( \lambda \)

We will work with an exponential distribution that has \( \lambda \) = 0.2
Therefore, the mean of this particular distribution is 0.5 (1/0.2)

## Simulation
We will simulate obtaining 1000 samples from an exponential distribution with \( \lambda \)=0.2, each of size 40 (n=40).  
So here we go:


```{r echo=TRUE}
averages <- NULL

for (i in 1:1000)
{ 
  averages <- data.frame( x = c(averages, mean(rexp(40, 0.2))))
}

str(averages)

```

The average of the 1000 sample means is
```{r echo=FALSE}
print(mean(averages))
```

This is almost identical to the theoretical mean for an exponential distribution with \( \lambda \) = 0.2

Our original standard deviation is the same as the mean, 1 / lambda, so we just have to square it to get the variance for the exponential distribution with \( \lambda \) = 0.2:

```{r echo=TRUE}
theoreticalVariance <- (1/0.2)^2 / 40
print(theoreticalVariance)
```

If we then calculate the variance of our own sampling data, we get a close approximation:

```{r echo=TRUE}
print(var(averages))
```


```{r}
ggplot(data = averages, aes(x = x)) + 
        geom_histogram(
                aes(y = ..density..),
                binwidth = 0.15,
                color = I("black"), 
                fill = I("blue")) + 
        stat_function(
                fun = dnorm, 
                arg = list(mean = mean(averages), sd = sqrt(theoreticalVariance)), 
                size = 1.2) + 
        labs(
                title = "Sampling Distribution of the Sample Means",
                x = "Sample Means",
                y = "Frequency")
