---
title: "Effect of Vitamin C on Tooth Growth of Guinea pigs"
author: "Tomás A. Maccor"
date: "22/3/2020"
output: pdf_document
---

```{r setup, include=FALSE}
```

## Introduction

This assignment explores whether 3 different doses & 2 different delivery methods of Vitamin C have an influence on the tooth lengths of Guinea pigs. The dataset used is **ToothGrowth**, which comes with R as one of its practice/sample datasets. 

Each animal received one of 3 doses of vitamin C, by one of 2 delivery methods: ascorbic acid (VC) versus orange juice (OJ).

This report comprises:

- A basic/descritive statistics summary of our sample data, plus an exploratory analysis.
- Use of statistical inference methods so that if any effects/conclusions are obtained from the SAMPLE data (contained in the ToothGrowth dataset), we can infer that these effects also apply to the entire population of Guinea pigs.


## Summary of data & exploratory analysis

First glimpse of the ToothGrowth dataset:

```{r echo=FALSE}
library(ggplot2, quietly = TRUE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)

## load Tooth Growth data

data("ToothGrowth")
str(ToothGrowth)
```

LEN = length of odontoblast.  NUMERIC variable   
SUPP = delivery method (VC or OJ).   FACTOR   
DOSE = Vitamin C dose: 0.5, 1 & 2 mg/day.    NUMERIC variable   


How many Guinea pigs received which dose, and by which method?:

```{r results = 'asis'}
a <- ToothGrowth %>% group_by(dose, supp) %>% summarise(n = n())
knitr::kable((a), caption = "Nr. of observations in each group")
```

So then  -20 pigs received each of the 3 doses ---> 10 via VC, 10 via OJ.

Let's have a first visualization of the data:

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

dose_supp_plot <- ggplot(ToothGrowth, aes(x = dose, y = len, fill = supp))
dose_supp_plot + geom_boxplot()
```

We can observe what could be a significant difference in Tooth Growth between the 0.5 mg/day dose
versus the 1 & 2 mg/day doses. 
We need to test for statistical significance to ascertain this.
It would also appear there is a difference between delivering Vit_C via *VC* vs *OJ* for the *0.5* and *1mg/day* doses.

If we plot the tooth growth differences by Vitamin C Dose only, and obtain the means for the 3 different Vit C doses:

```{r}
dose_plot <- ggplot(ToothGrowth, aes(x = dose, y = len))
dose_plot + geom_boxplot()

b <- ToothGrowth %>% group_by(dose) %>% summarise(Mean_Tooth_Growth = mean(len))
knitr::kable(b)
```

This time, it could well be that there's statistical difference between all 3 doses.

As we have > 2 groups with what appear similar intergroup variance, we use ANOVA for testing significance (equal variance).

So far, our assumptions are:
- That tooth growth in Guinea pigs is normally distributed
- That the 3 different DOSE groups have an equal variance

```{r}
ANOVA_dose <- aov(len ~ dose, data = ToothGrowth)
summary.aov(ANOVA_dose)
```

So, per ANOVA, we have a statistically significant difference in Mean Tooth Growth between the 3 DOSES (p = 9.53e-16, thus p < 0.05).

But, ANOVA does not allow to know which of the pairwise DOSE comparisons are significant  --so now we perform TUKEY TEST to determine this:


```{r}
TukeyHSD(ANOVA_dose)
```

Which also results in statistically significant differences between all 3 doses (p < 0.05).
The *differences between the mean DOSES* and their *confidence intervals for those mean differences* are listed(provided) in the Tukey Test.

So for example, we can state that if the entire population of Guinea pigs was given Vitamin C at 3 doses, and we took random samples of these pigs, 95% of the times we would obtain a Mean difference in Tooth Growth that would be between 5.90 to 12.36 (in the 0.5 mg vs. 1.0 mg/day Vitamin C groups). 


In relation to the TYPE of Vitamin C supplement, it is not advisable to test solely between VC vs OJ, as DOSE could a confounder in this relation.

So we do pairwise t-tests for each dose

```{r}
ToothGrowth <- ToothGrowth %>% arrange(dose)
Dose0.5 <- ToothGrowth[1:20, ]
Dose1 <- ToothGrowth[21:40, ]
Dose2 <- ToothGrowth[41:60, ]

t.test(len ~ supp, data = Dose0.5)
t.test(len ~ supp, data = Dose1)
t.test(len ~ supp, data = Dose2)

```


